Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-05-27 3dpas_intro.tex
merged in YS changes:
replaced <CR> with <LF>
affiliation
added citations
modifactions mainly to infrastructure section

% cpedoc.tex V2.0, 13 May 2010

\documentclass[times]{cpeauth}

\usepackage{moreverb}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2012} %\documentclass[acmjacm,acmnow]{acmtrans2m}
%\documentclass[conference,final]{IEEEtran}
%\documentclass[10pt,letterpaper]{article}
%\documentclass[preprint,12pt]{article}
%\documentclass[11pt,letterpaper]{article}

\usepackage{graphicx} \usepackage{url} \usepackage{color} \usepackage{enumerate}
\usepackage{longtable} %\usepackage{authblk} % Is used internally within the CPE
Class \usepackage[firstpage]{draftwatermark}

%\SetWatermarkLightness{0.7} %\SetWatermarkScale{4}

%\usepackage{pdfsync} % \acmVolume{} % \acmNumber{} % \acmYear{} % \acmMonth{}

%\usepackage{sectsty} %\usepackage{setspace}

% \usepackage{ifpdf} % \usepackage{wrapfig} % %\usepackage[cm]{fullpage} %
\usepackage{textcomp} % \usepackage{srcltx} % \usepackage{fancyhdr} %
\usepackage{setspace} \usepackage{lscape} % \usepackage{longtable} %
\usepackage{paralist}

%\usepackage{hyperref} \usepackage[novbox]{pdfsync}

% \usepackage{ams}

% \ifpdf % \usepackage[pdftex, % colorlinks=true, % linkcolor=red, %
citecolor=red, % filecolor=red, % ]{hyperref} % \else %
\usepackage[hypertex]{hyperref} % \fi

% Space saving %\usepackage[small,compact]{titlesec}
%\usepackage[small,it]{caption}

% \renewcommand\floatpagefraction{.9} % \renewcommand\topfraction{.9} %
\renewcommand\bottomfraction{.9} %\renewcommand\textfrare{.1}

\setcounter{totalnumber}{50} \setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

%\textwidth = 6in %\textheight = 8.5 in %\oddsidemargin = 0.0 in
%\evensidemargin = 0.0 in

% \topmargin = 0.0 in % \headheight = 0.0 in % \headsep = 0.0 in
%\parskip = 0.0in
%\parindent = 0.5cm
\parskip = 0.04in
% \parindent = 0.0cm % \textfloatsep = 0.1in

% \usepackage[draft,pdftex,colorlinks=true, linkcolor=blue,citecolor=blue, %
urlcolor=blue]{hyperref}

% \ifpdf % \pdfinfo{ /Author (Shantenu Jha et al.)  /Title (Programming
Abstractions for Large-Scale Distributed Applications) } % \fi

%\newcommand{\projectnamefull}{\textit{Programming Abstractions for %
Large-Scale Distributed Applications} }

\newcommand{\upup}{\vspace*{-0.5em}} \newcommand{\up}{\vspace*{-0.25em}}
\newcommand{\I}[1]{\textit{#1}} \newcommand{\B}[1]{\textbf{#1}}
\newcommand{\T}[1]{\texttt{#1}} \newcommand{\BI}[1]{\B{\I{#1}}}

 \newenvironment{shortlist}{ \vspace*{-0.8em}
  \begin{itemize} \setlength{\itemsep}{-0.3em} }{
  \end{itemize} \vspace*{-0.6em} }

 \newenvironment{shortenum}{ \vspace*{-0.8em}
  \begin{enumerate} \setlength{\itemsep}{-0.3em} }{
  \end{enumerate} \vspace*{-0.6em} }

% \ifpdf % \DeclareGraphicsExtensions{.pdf, .png, .jpg} % \else %
\DeclareGraphicsExtensions{.ps, .eps} % \fi

\definecolor{orange}{rgb}{1.0,0.3,0.0} \definecolor{violet}{rgb}{0.75,0,1}
\definecolor{darkgreen}{rgb}{0,0.6,0} \definecolor{cyan}{rgb}{0.2,0.7,0.7}
\definecolor{blueish}{rgb}{0.2,0.2,0.8}

\long\def\comment#1{{\bf \textcolor{magenta}{\bf #1}}} \long\def\ccomment#1{{\bf
\textcolor{blue}{\bf #1}}} \newcommand{\C}{\comment} \newcommand{\CC}{\ccomment}

\newcommand{\yes}{$\bullet$}

\long\def\finalcomment#1{{\bf \textcolor{red}{\bf #1}}}
\newcommand{\CF}{\finalcomment}

\newif\ifdraft \drafttrue \ifdraft \newcommand{\katznote}[1]{
{\textcolor{magenta} { ***Dan: #1 }}} \newcommand{\rananote}[1]{
{\textcolor{cyan} { ***Omer: #1 }}} \newcommand{\omernote}[1]{ {\textcolor{cyan}
{ ***Omer: #1 }}} \newcommand{\note}[1]{ {\textcolor{red} { #1 }}}
\newcommand{\jhanote}[1]{ {\textcolor{blue} { ***Shantenu: #1 }}}
\newcommand{\jonnote}[1]{ {\textcolor{red} { ***Jon: #1 }}}
\newcommand{\alnote}[1]{ {\textcolor{darkgreen} { ***Andre: #1 }}}
\newcommand{\nchnote}[1]{ {\textcolor{orange} { ***Neil: #1 }}}
\newcommand{\ysnote}[1]{ {\textcolor{violet} { ***Yogesh: #1 }}}
\newcommand{\sdnote}[1]{ {\textcolor{blueish} { ***Simon: #1 }}} %
\usepackage[pdftex,colorlinks=true, linkcolor=blue,citecolor=blue, %
urlcolor=blue]{hyperref} \else \newcommand{\katznote}[1]{}
\newcommand{\rananote}[1]{} \newcommand{\note}[1]{} \newcommand{\jhanote}[1]{}
\newcommand{\jonnote}[1]{} \newcommand{\alnote}[1]{} \newcommand{\nchnote}[1]{}
\newcommand{\ysnote}[1]{} \fi

@@

%We always run on all available resources, so if we have computer time %across
the country, then our data will be distributed, and so will our
%computation. ITER will have data that will be distributed, but it %still
unknown.

@@

%I don't understand this question? We will input data (from perhaps a restart),
and this data is typically in the ADIOS-BP file format for large data. Smaller
data sets are typically stored in HDF5, and the smallest are stored in MDS+.

@@

%Depends on what we want to understand --- For GTS/GTC we have typically %looked
at the particle data to compute fluxes, etc. and this data can %be large, but
due to the Monte Carlo error, and ways to generated 5D %distribution functions
from the particles, we reduce the particles to %a manageable amount (300 GB per
time step x 1 K time steps). Our %current I/O is capable of reading this in $<$
100 s per time step, which %allows us to analyze the data in about 1-2 days.

@@

%Don't understand the question. Features change and are dynamic. The physics
will get more complex as we continue to evolve our codes, and will be more
dynamic in nature.

@@

%Reading the data from disk --- (Perhaps I don't understand the question).  It
will also couple to other codes, and the coupling will be in-memory data
exchange.
%6. Are there time (or quality) constraints on the application?  (If so, what
are they?)
%We always need to run as fast as possible since we have limited computing time
(our team does have about 100M hours FY 11, but this still puts severe
restrictions on what we run, so the faster we can run, the more science we can
get done.
%7. Please feel free to also talk about the current state of the application, if
it exists today, and any specific gaps that you know will need to be overcome.
%Our biggest gaps will be in accelerating the code on next generation exascale
platforms, and in processing as much data as possible in situ.

@@

The application~\cite{modisAzure,modis:ipdps:2010} is used to handle the processes of downloading
data from different sites and integrating them. The integration consists of
reprojecting the data from one map projection to another, spatial and temporal
resampling, and gap filling. The application itself is a loosely-coupled data
pipeline that coordinates these processes.

@@

{\em Data Sources} are used in the acquisition stage as inputs of data into the
application. Broadly speaking these sources include types such as files, servers,
sensors, instruments, simulations and models.  Additionally, databases are often
considered as an intermediate data source in types of system. The dynamic
characteristics are primarily due to availability. These are often accessed in
the Data Generation stage of the application
(\figurename~\ref{fig:figures_application-stages}).

{\em Data Storage} components are used to provide persistent access to data by
the application. They may be structured (e.g., databases, No-SQL tables) or
unstructured (e.g., collections of audio recordings). The dynamic
characteristics are primarily due to the level of persistence, e.g., temporary,
durable, permanent (archived or published), or replicated; and also in the
structure of the data stored, e.g., schema free tables as opposed to relational
tables. There may also be dynamic characteristics of storage due to rate of
growth of data, e.g., a field survey may bring in a lot of data during summer
but none during other times.  These are commonly used in the Data Storage stage
of the application (\figurename~\ref{fig:figures_application-stages}), but may
also be present at the boundaries between the stages for transient staging of
data.

@@

{\em Data Movement} components are used for bulk or continuous movement of data
between parts of the application, or to and from the application. % (often
called staging) Data movement may be explicit and visible to the application
(e.g., staging), or implicit (e.g., automatic replication of files). Components
may operate on streams or blocks (files). Dynamic characteristics include
partitioning, scheduling and planning (e.g., edge caching of ``hot'' data, where
datasets that are being used frequently during some period are replicated close
to the application). Data movement can be considered to be real-time (i.e., as
it happens, e.g., sensor data), as real-time (e.g., a storage buffer has been
introduced into the transport system, but the data source is still real-time, as
with time-series data), or asynchronous (where the ordering of the data movement
is independent of the data itself).  Latency may be considered as an important
characteristic of the data movement infrastructure depending on the application
e.g. real-time processing of sensor data for disaster analysis. Other dynamic
aspects of data movement include data rates, their variability from data
sources, and the ``freshness'' of data. These infrastructure appear on the edges
between the application stages.

{\em Data Discovery} components assist the access to and use of data sources,
and include Catalogs, Information Services, and Metadata stores. The dynamic
characteristics pertain to the evolution or recalibration of information held in
these components. These components are orthogonal to the applications stages and
can be used by any of the stages to guide their selection or identification of
incoming data sources and outgoing data sinks.  %\ysnote{rate of growth of data
may be an issue.  %Catalogs that perform nightly ``crawling'' or data loading
(e.g., sky surveys) may not be suitable if the %latest data needs to be
accessible to applications. Catalog update will have to keep up with data
%update. However, batch processing applications can survive less frequent
updates.}

@@

Analysis components include functionality such as comparison, statistical
analysis, event analysis (i.e., analyzing information derived from data rather
than raw data) and visualization. These components leverage the computational
aspects of the infrastructure.  Based on the application stages outlined in
\figurename~\ref{fig:figures_application-stages}, we identify a number of
infrastructure services that can be associated with these stages.

@@

%\ysnote{addresses Andre's commment} {\em Analytics and visualization}
components help extract domain knowledge and assist with exploration of the
data. These are often closely supervised by the end-user and may even be
interactive. The proliferation of ``big data'' has caused data mining libraries,
machine-learning toolboxes and visualization environments to become part of the
infrastructure, some of which are deployed for specific domains but others which
are more generic (e.g., Weka, GraphLab). These components are associated with
the Data Processing stage.

@@

{\em Orchestration} components handle the management and choreography of data
placement, code placement, communication, and messaging to allow complex
analyses to be performed on distributed data sources. They may also be dynamic
if they have the capability to adapt their behavior based on earlier parts of
the workflow or can be steered based on external notification received during
execution. In addition to supporting choreography, often making use of one or
more resource manager/scheduler components, an orchestration component may also
include specialist support for conversion and analysis components discussed
above.

{\em Bulk Data-parallel} frameworks use ``Bulk'' operators to express
computations over collections of data which are executed in a loosely-coupled,
data-parallel manner.  The {\em MapReduce} paradigm~\cite{MapReduce} and its
variants~\cite{twister,hmr} are the most prominent among these. MapReduce forces
programmers to model computation in two stages: a \emph{map} phase in which a
transformation is applied to every element of a collection independently of
every other; and a \emph{reduce} phase in which the transformed results are
aggregated through a binary function that is associative and commutative.  These
properties serve to guarantee that the MapReduce computation can be efficiently
parallelized, by performing the map phase concurrently and then reducing without
the reduction order affecting the final result.

@@

\hline \multicolumn{2}{|l|}{{\textbf{Application Name}}} \\ \hline %\nchnote{For
example, workload decomposition and distribution must be determined dynamically
in order to optimally use resources, e.g., compute resource selection can be
based on data location, processing profile and/or network capacity. Workflows?
Resource selection?}  Infrastructure for Dynamic Properties & Infrastructure for
Distributed Properties \\ \hline
%\begin{footnotesize}
% \begin{longtable}[h!]{|p{6.7cm}|p{6.7cm}|}
% \caption{Infrastructure used to support dynamic and distributed properties of
application scenarios} \label{tab:inf_app_mapping} \\ % \hline %
\textbf{Infrastructure to support Dynamism} % &\textbf{Infrastructure to support
Distributedness} \\
% \endhead \hline \multicolumn{2}{|l|}{{\textbf{Next Generation Sequencing (NGS)
Analytics}}} \\ \hline %\nchnote{For example, workload decomposition and
distribution must be determined dynamically in order to optimally use resources,
e.g., compute resource selection can be based on data location, processing
profile and/or network capacity. Workflows? Resource selection?}  Dynamic
workload decomposition and distribution for resource optimization
\emph{(Orchestration)} & Distributing data from source (generation) to the
computing destination (analysis) \emph{(Data Discovery, Movement)} % No unifying
single requirement or single solution at present (or possible at all?)  %
Normally data local and CI doesn't easily support movement of data to compute %
\ysnote{We may want to remove the preceding lines since we're talking about
existing infrastructure, not missing ones at this point.}  % naturally
distributed: data source (NGS machine storage) and compute on % different
resources (typically not co-located); loosely coupled (many % machines,
different repositories) \\ \hline \multicolumn{2}{|l|}{{\textbf{ATLAS (an
example of use of the WLCG) Reconstruction app}}} \\ \hline Manage gradual
growth of data using SRM \emph{(Data Storage, Access)} % Management, subset
replication and partitioning to go inf components: data replication, (gLite),
data movement % specify SRM as standard and then they leave choice of
implementation to the organisations as part of consortium.  % New raw data is
collected % steadily (2) \note{timescale needs to worked out. What happens if
something is found?}  & Management, subset replication, data movement and
partitioning using gLite and SRM \emph{(Data Storage, Access, Movement,
Discovery)} % , data movement % natural, (sub-set) replicated, loosely coupled
\note{Is natural the % right word? Artificial % distribution (the source is
natural, but the intermediate points are % artificial), data not
partitioned}\ysnote{These details are not fully present in the actual
application description of its dynamism (Sec 2.2). Should we trim there and move
some content there? } \\ \hline \multicolumn{2}{|l|}{{\textbf{LSST}}} \\ \hline

Steering observations, Databases to manage data growth \emph{(Notification, Data
Storage)} % DE decisions based on observed data (2, 5) \ysnote{what do these
numbers mean? specific % characteristics?}  & Coordinate classification,
distributed databases, data movement to distributed observatories
\emph{(Notification, Data Access, Data Movement, Data Storage)} %Analysis, % ,
natural distribution of telescopes \\ \hline \multicolumn{2}{|l|}{{\textbf{SOA
Astronomy Applications}}} \\ \hline

Catalog contents updated, Query mix changes \emph{(Data Discovery)} % data is
continuously updated, thus dynamic data (pre-defined schedules or on-demand),
queries change (7)\note{timescale clarification!}  & Distributed VO Services for
data storage and access \emph{(Data Discovery, Data Access, Data Movement)} %
moving data sources, services and storage are distributed across institutions \\
\hline \multicolumn{2}{|l|}{{\textbf{Cosmic Microwave Background}}} \\ \hline
Dynamic resource selection %dynamic utilization of resources (4) & Compute on
distributed HPC centers \emph{(Data Movement)} % (sub-set) replication (during
application run) \note{What phases of the application do we consider? data
acquisition? app run?}\ysnote{Can any of the supercomputers be selected? If so,
the data is moved to distributed compute (or it is dynamic resource selection)}
\\ \hline \multicolumn{2}{|l|}{{\textbf{Sensor Network Application}}} \\ \hline
Notification of sensor data when in range, data rates and volumes \emph{(Data
Movement, Notification)} %data rate, volume dynamic (1-2) &Aggregating
distributed sensor data \emph{(Data Movement, Notification)} \\ \hline
\multicolumn{2}{|l|}{{\textbf{Climate (ESGF)}}} \\ \hline Manage data growth %
New data generated over time (ESGF-wide - is this in scope of the application?)
& Distributed, replicated catalogs using ESGF software stack (GridFTP,
wget/HTTP) \emph{(Data Access, Data Discovery, Data Movement, Data Conversion)}
\\ % multiple climate centers (natural), users can request data from ESGF
(sub-set replication), coupled % Can come in via website - cookie / OpenID %
Come in through command line script - grid cert \hline
\multicolumn{2}{|l|}{{\textbf{Interactive Exploration of Environmental Data}}}
\\ \hline On-demand data subsetting, caching and versioning, real-time
instrument data \emph{(Data Discovery, Data Movement, Data Sources)} % at moment
services are hosted on a single address. There may be farming for load balancing
in the future.  % data rates change (1), % data inputs/outputs change (3),
queries change (7) & Distributed Data Services, Access Restrictions \emph{(Data
Movement, Data Access)}

@@

\\ \hline % Dynamic Environmental Data on the Web &data operated on changes %
(realtime feeds) (2) % &natural (sensor sources, simulations), (sub-set)
replication\\ % \hline \multicolumn{2}{|l|}{{\textbf{PowerGrids}}} \\ \hline

@@

		\hline \multicolumn{2}{|l|}{{\textbf{Fusion (ITER)}}} \\ \hline
Stream and transform data between simulation sites on-demand \emph{(Data
Movement, Enrichment)} % (multiple distinct data % flows), on-the-fly %
transformation (8) & Run analysis on available, distributed resources
\emph{(Data Movement)} %Analysis % natural, coupled, transient sub-set
replicated \\ \hline \multicolumn{2}{|l|}{{\textbf{Industrial Incident
Notification and Response}}} \\ \hline Acquire different data sources and rates,
Processing determined by data \emph{(Data Sources, Data Access, Notification,
Orchestration)} %Analysis, % (1), % data itself changes (2), dynamic data
sources (3), processing changes % based on data %(5) & Distributed data sources,
Analysis by agents across institutions \emph{(Data Sources, Data Access)} % ,
Analysis % natural, coupled \\ \hline \multicolumn{2}{|l|}{{\textbf{MODIS Data
Processing}}} \\ \hline

@@

\multicolumn{2}{|l|}{{\textbf{Distributed Network Intrusion Detection}}} \\
\hline Analyze dynamic network traffic rates \emph{(Data Movement)} % , Analysis
% data rate variable (1), data % highly dynamic changing (2), application
responds to data (5), % on-the-flight processing (8) & Run analysis at different
sites, Distribute generated data \emph{(Data Movement)} % Analysis, % natural,
data/compute tightly coupled \\ \hline
% \end{longtable}
%\end{footnotesize}

@@

\begin{table}
  \begin{scriptsize}
    \begin{center}
      \caption{Coordination approaches used by the
applications\label{table:progSystems}}
      \begin{tabular}{|p{6.3cm}|p{7.2cm}|} \hline Application & Current
approaches \\ \hline \hline NGS Analytics & pipelines that ship with
sequencers~\cite{illumina}; workflows~\cite{Taverna} \\ \hline ATLAS & custom
algorithms; job submission tied to WLCG \\ \hline LSST & custom layer; scripted
workflows; queries \\ \hline SOA Astronomy & web services \\ \hline Cosmic
Microwave Background & \textit{ad hoc} \\ \hline Sensor Network Application &
focused on data collection \\ \hline Climate & \textit{ad hoc} \\ \hline
Interactive Exploration of Environmental Data & visualization tools \\ \hline
Power Grids & stream processing; data mining \\ \hline Fusion (ITER) & scripted
pipeline \\ \hline Industrial Incident Notification and Response & agent-based
\\ \hline MODIS Data Processing & scripted pipeline \\ % \hline Floating Sensors
& \textit{ad hoc} \\ \hline Distributed Network Intrusion Detection & monolithic
packages \\ \hline
      \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

>>>>

\include{header}

@@

%We always run on all available resources, so if we have computer time
%across the country, then our data will be distributed, and so will our
%computation. ITER will have data that will be distributed, but it
%still unknown.

@@

%I don't understand this question? We will input data (from perhaps a restart), and this data is typically in the ADIOS-BP file format for large data. Smaller data sets are typically stored in HDF5, and the smallest are stored in MDS+.

@@

%Depends on what we want to understand --- For GTS/GTC we have typically
%looked at the particle data to compute fluxes, etc. and this data can
%be large, but due to the Monte Carlo error, and ways to generated 5D
%distribution functions from the particles, we reduce the particles to
%a manageable amount (300 GB per time step x 1 K time steps). Our
%current I/O is capable of reading this in $<$ 100 s per time step, which
%allows us to analyze the data in about 1-2 days.

@@

%Don't understand the question. Features change and are dynamic. The physics will get more complex as we continue to evolve our codes, and will be more dynamic in nature.

@@

%Reading the data from disk --- (Perhaps I don't understand the question).  It will also couple to other codes, and the coupling will be in-memory data exchange.
%6. Are there time (or quality) constraints on the application?  (If so, what are they?)
%We always need to run as fast as possible since we have limited computing time (our team does have about 100M hours FY 11, but this still puts severe restrictions on what we run, so the faster we can run, the more science we can get done.
%7. Please feel free to also talk about the current state of the application, if it exists today, and any specific gaps that you know will need to be overcome.
%Our biggest gaps will be in accelerating the code on next generation exascale platforms, and in processing as much data as possible in situ.

@@

The application~\cite{modisAzure, modis:ipdps:2010} is used to handle the processes of downloading data from different
sites and integrating them. The integration consists of reprojecting the data from one map projection to another,
spatial and temporal resampling, and gap filling. The application itself is a loosely-coupled data
pipeline that coordinates these processes.

@@

{\em Data Sources} are used in the acquisition stage as inputs of data
into the application. Broadly speaking these sources include types such as
files, servers, sensors, instruments, simulations and models.
Additionally, databases are often considered as an intermediate data
source in some systems. The dynamic characteristics are primarily
due to availability. These are often accessed in the Data Generation stage of the application
(\figurename~\ref{fig:figures_application-stages}).

{\em Data Storage} components are used to provide persistent access to
data by the application. They may be structured (e.g., databases,
No-SQL tables) or unstructured (e.g., collections of audio
recordings). The dynamic characteristics are primarily due to the
level of persistence, e.g., temporary, durable, permanent (archived or
published), or replicated; and also in the structure of the data
stored, e.g., schema free tables as opposed to relational tables. There
may also be dynamic characteristics of storage due to rate of growth
of data, e.g., a field survey may bring in a lot of data during summer
but none during other times. Some storage components like files and databases may also serve as data
sources. These are commonly used in the Data Storage stage of the application (\figurename~\ref{fig:figures_application-stages}),
but may also be present at the boundaries between the stages for transient staging of data.

@@

{\em Data Movement} components are used for bulk or continuous
movement of data between parts of the application, or to and from the
application, % (often called staging) Data movement may be explicit
and visible to the application (e.g., staging), or implicit (e.g.,
automatic replication of files). Components may operate on streams or
blocks (files). Dynamic characteristics include partitioning,
scheduling and planning (e.g., edge caching of ``hot'' data, where
datasets that are being used frequently during some period are
replicated close to the application). Data movement can be considered
to be real-time (i.e., as it happens, e.g., sensor data), `as real-time'
(e.g., a storage buffer has been introduced into the transport system,
but the data source is still real-time, as with time-series data), or
asynchronous (where the ordering of the data movement is independent
of the data itself).  Latency may be considered as an important
characteristic of the data movement infrastructure depending on the
application, e.g., real-time processing of sensor data for disaster
analysis. Other dynamic aspects of data movement include data rates,
their variability from data sources, and the ``freshness'' of
data. These infrastructure appear on the edges between the application
stages.

{\em Data Discovery} components assist the access to and use of data
sources, and include Catalogs, Information Services, and Metadata
stores. The dynamic characteristics pertain to the evolution or
recalibration of information held in these components. These components are orthogonal to the
applications stages and can be used by any of the
stages to guide their selection or identification of incoming data sources and outgoing data sinks
or storage.
%\ysnote{rate of growth of data may be an issue.
%Catalogs that perform nightly ``crawling'' or data loading (e.g., sky surveys) may not be suitable if the
%latest data needs to be accessible to applications. Catalog update will have to keep up with data
%update. However, batch processing applications can survive less frequent updates.}

@@

Analysis components include functionality such as comparison,
statistical analysis, event analysis (i.e., analyzing information
derived from data rather than raw data) and visualization. These components leverage
the computational aspects of the infrastructure.
Based on the application stages outlined in
\figurename~\ref{fig:figures_application-stages}, we identify a number
of infrastructure services for data analysis that can be associated with these stages.

@@

%\ysnote{addresses Andre's commment}
{\em Analytics and visualization} components help extract domain knowledge and assist with
exploration of the data. These are often closely supervised by the end-user and may even be
interactive. The proliferation of ``big data'' has caused data mining libraries, machine-learning
toolboxes and visualization environments to become part of the infrastructure, some of which are
deployed for specific domains but others which are more generic (e.g., Weka, GraphLab, Apache Mahout). These components are associated
with the Data Processing stage.

@@

{\em Orchestration} components handle the management and
choreography of data placement, code placement, communication, and
messaging to allow complex analyses to be performed on distributed
data sources. These are often composed as workflows or Directed Acyclic Graphs (DAGs) that capture
the execution dependency between
components. They may also be dynamic if they have the capability to
adapt their behavior based on earlier parts of the workflow or can be
steered based on external notification received during execution. In
addition to supporting choreography, while often making use of one or more
resource manager/scheduler components, an orchestration component may also
include specialist support for conversion and analysis components
discussed above.

{\em Bulk Data-parallel} frameworks use ``Bulk'' operators to express computations over 
collections of data which are executed in a loosely-coupled, data-parallel manner.
The {\em MapReduce} paradigm~\cite{MapReduce} and its variants~\cite{twister,hmr} are the most
prominent among these, with recent ones like Pregel and Apace Giraph~\cite{pregel} supporting specific domains like
graph analytics. MapReduce forces programmers to model computation in two stages: 
a \emph{map} phase in which a transformation is applied to every element of a collection independently of every other;  and a \emph{reduce} phase in which the transformed results 
are aggregated through a binary function that is associative and commutative. 
These properties serve to guarantee that the MapReduce computation can be efficiently 
parallelized, by performing the map phase concurrently and
then reducing without the reduction order affecting the final result.

@@

\hline	
\multicolumn{2}{|l|}{{\textbf{Application Name}}} \\ \hline	
%\nchnote{For example, workload decomposition and distribution must be determined dynamically in order to optimally use resources, e.g., compute resource selection can be based on data location, processing profile and/or network capacity. Workflows? Resource selection?}
Infrastructure to support Dynamic Properties & Infrastructure to support Distributed Properties
 \\ \hline
%\begin{footnotesize}
%	 \begin{longtable}[h!]{|p{6.7cm}|p{6.7cm}|}
%		\caption{Infrastructure used to support dynamic and distributed properties of application scenarios} \label{tab:inf_app_mapping} \\
%		\hline
%		\textbf{Infrastructure to support Dynamism}
%                &\textbf{Infrastructure to support Distributedness} \\		
%		 \endhead
		\hline	
\multicolumn{2}{|l|}{{\textbf{Next Generation Sequencing (NGS) Analytics} (\S\ref{bioSilvia})}} \\ \hline	
%\nchnote{For example, workload decomposition and distribution must be determined dynamically in order to optimally use resources, e.g., compute resource selection can be based on data location, processing profile and/or network capacity. Workflows? Resource selection?}
Dynamic workload decomposition and distribution for resource optimization \emph{(Orchestration)}
		&
Distributing data from source (generation) to the computing destination (analysis) \emph{(Data
  Discovery, Movement)}
% 		No unifying single requirement or single solution at present (or possible at all?)
% 		Normally data local and CI doesn't easily support movement of data to compute
% \ysnote{We may want to remove the preceding lines since we're talking about existing infrastructure, not missing ones at this point.}
% naturally distributed: data source (NGS machine storage) and compute on
% 		different resources (typically not co-located); loosely coupled (many
%		machines, different repositories)
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{ATLAS (an example of use of the WLCG) Reconstruction app} (\S\ref{WLCGSteve})}} \\ \hline
Manage gradual growth of data using SRM \emph{(Data Storage, Access)}
% Management, subset replication and partitioning to go inf components: data replication, (gLite), data movement
% specify SRM as standard and then they leave choice of implementation to the organisations as part of consortium.
% New raw data is collected
% 	    steadily (2) \note{timescale needs to worked out. What happens if something is found?}
&
Management, subset replication, data movement and partitioning using gLite and SRM \emph{(Data
  Storage, Access, Movement, Discovery)}
% , data movement
% natural, (sub-set) replicated, loosely coupled \note{Is natural the
% 	   right word? Artificial
% 		distribution (the source is natural, but the intermediate points are
% 		artificial), data not partitioned}\ysnote{These details are not fully present in the actual application description of its dynamism (Sec 2.2). Should we trim there and move some content there? }
              \\
		\hline
\multicolumn{2}{|l|}{{\textbf{LSST} (\S\ref{astroAdam})}} \\ \hline

Steering observations, Databases to manage data growth \emph{(Notification, Data Storage)}
% DE decisions based on observed data (2, 5) \ysnote{what do these numbers mean? specific
             %  characteristics?}
            &
Coordinate classification, distributed databases, data movement to distributed observatories
\emph{(Notification, Data Access, Data Movement, Data Storage)} %Analysis, % , natural distribution of telescopes
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{SOA Astronomy Applications} (\S\ref{astroSOAAdam})}} \\ \hline

Catalog contents updated, Query mix changes \emph{(Data Discovery)}
% data is continuously updated, thus dynamic data (pre-defined schedules or on-demand), queries change (7)\note{timescale clarification!}
             &
Distributed VO Services for data storage and access \emph{(Data Discovery, Data Access, Data Movement)}
% moving data sources, services and storage are distributed across institutions
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{Cosmic Microwave Background} (\S\ref{astroJulian})}} \\ \hline	
	      Dynamic resource selection  \emph{(Data Discovery)}
%dynamic utilization of resources (4)
& Compute on distributed HPC centers \emph{(Data Movement)}
% (sub-set) replication (during application run) \note{What phases of the application do we consider? data acquisition? app run?}\ysnote{Can any of the supercomputers be selected? If so, the data is moved to distributed compute (or it is dynamic resource selection)}
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{Sensor Network Application} (\S\ref{sensorSimon})}} \\ \hline	
	     Notification of sensor data when in range, data rates and volumes \emph{(Data Movement, Notification)} %data rate, volume dynamic (1-2)
&Aggregating distributed sensor data \emph{(Data Movement, Notification)}  \\
		\hline
\multicolumn{2}{|l|}{{\textbf{Climate (ESGF)} (\S\ref{climateDan})}} \\ \hline	
	      Manage data growth % New data generated over time (ESGF-wide - is this in scope of the application?)
             &
Distributed, replicated catalogs using ESGF software stack (GridFTP, wget/HTTP) \emph{(Data Access,
  Data Discovery, Data Movement, Data Conversion)} \\
  % multiple climate centers (natural), users can request data from ESGF (sub-set replication), coupled
% Can come in via website - cookie / OpenID
% Come in through command line script - grid cert
		\hline
\multicolumn{2}{|l|}{{\textbf{Interactive Exploration of Environmental Data} (\S\ref{envJon})}} \\ \hline	
On-demand data subsetting, caching and versioning, real-time instrument data \emph{(Data Discovery, Data
  Movement, Data Sources)}
% at moment services are hosted on a single address. There may be farming for load balancing in the future.
% data rates change (1),
% 	    data inputs/outputs change (3), queries change (7)
		&
Distributed Data Services, Access Restrictions \emph{(Data
  Movement, Data Access)}

@@

\\
		\hline
	    % Dynamic Environmental Data on the Web &data operated on changes
	    %     (realtime feeds) (2)
	    %     &natural (sensor sources, simulations), (sub-set) replication\\
	    %     \hline
\multicolumn{2}{|l|}{{\textbf{PowerGrids} (\S\ref{power})}} \\ \hline

@@

		\hline
\multicolumn{2}{|l|}{{\textbf{Fusion (ITER)} (\S\ref{fusionScott})}} \\ \hline	
	     Stream and transform data between simulation sites on-demand \emph{(Data Movement, Enrichment)}
%  (multiple distinct data
	    % flows), on-the-fly
	    % transformation (8)
	    &
Run analysis on available, distributed resources \emph{(Data Movement)} %Analysis
% natural, coupled, transient sub-set replicated
            \\
		\hline
\multicolumn{2}{|l|}{{\textbf{Industrial Incident Notification and Response} (\S\ref{Kees})}} \\ \hline	
	     Acquire different data sources and rates, Processing determined by data
             \emph{(Data Sources, Data Access, Notification, Orchestration)}   %Analysis,  % (1),
	    % data itself changes (2), dynamic data sources (3), processing changes
	    % based on data
            %(5)
&
Distributed data sources, Analysis by agents across institutions, Access control \emph{(Data Sources, Data
  Access)} % , Analysis
% natural, coupled
            \\
		\hline
\multicolumn{2}{|l|}{{\textbf{MODIS Data Processing} (\S\ref{modisKeith})}} \\ \hline

@@

\multicolumn{2}{|l|}{{\textbf{Distributed Network Intrusion Detection} (\S\ref{detectionJon})}} \\ \hline	
Analyze dynamic network traffic rates \emph{(Data Movement)} % , Analysis
	    %  data rate variable (1), data
	    % highly dynamic changing (2), application responds to data (5),
	    % on-the-flight processing (8)
            &
Run analysis at different sites, Distribute generated data \emph{(Data Movement)} % Analysis,
% natural, data/compute tightly coupled
\\
		\hline
%	 \end{longtable}
%\end{footnotesize}

@@

\begin{table}
  \begin{scriptsize}
    \begin{center}
      \caption{Coordination approaches used by the applications\label{table:progSystems}}
      \begin{tabular}{|p{6.3cm}|p{7.2cm}|}
        \hline Application  & Current approaches \\
        \hline
        \hline NGS Analytics & pipelines that ship with
        sequencers~\cite{illumina}; workflows~\cite{Taverna} \\
        \hline ATLAS & custom algorithms; job submission tied to WLCG \\
        \hline LSST & custom layer; scripted workflows; queries \\
        \hline SOA Astronomy & web services~\cite{barker} \\
        \hline Cosmic Microwave Background & \textit{ad hoc} \\
        \hline Sensor Network Application & focused on data collection \\
        \hline Climate & \textit{ad hoc} \\
        \hline Interactive Exploration of Environmental Data &
        visualization tools~\cite{blower2} \\
        \hline Power Grids & stream processing~\cite{Simmhan:sciencecloud:2011}; data mining~\cite{Yin:mapreduce:2012} \\
        \hline Fusion (ITER) & scripted pipeline~\cite{adios} \\
        \hline Industrial Incident Notification and Response &
        agent-based~\cite{emergency4}  \\
        \hline MODIS Data Processing & scripted pipeline~\cite{modis:ipdps:2010} \\
%        \hline Floating Sensors & \textit{ad hoc} \\
        \hline Distributed Network Intrusion Detection & monolithic
        packages \\
        \hline
      \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}
